---
title: "01 - Inspection of data"
author: "Jens Linden"
output:
  html_document:
    number_sections: yes
    toc: yes
---

STATUS: Development

# Management summary

* Load raw data
* Inspect data and perform initial data checks
* Preprocessing data
* Saving preprocessed data to file 
    + Preprocessed review data table
    + Hotel master data table
    + Tag table
    + Table for tag statistics

The following hypotheses have been tested:

| ID | Hypothesis | Result|
|------|-----------------------------------|-----------------------------------|
| H01-01 | The data provided for the challenge is identical to the [Kaggle 515K Hotel Reviews Data in Europe](https://www.kaggle.com/jiashenliu/515k-hotel-reviews-data-in-europe) | Yes |
| H01-02 | Review_Date is consistently in format mm/dd/yyyy | Yes |
| H01-03 | Geolocations are available for all hotels | No, missing for 17 hotels |
| H01-04 | The hotels are from all over Europe | No, the hotels are from 6 cities only (London, Amsterdam, Paris, Milan, Barcelona, Vienna) |
| H01-05 | All free text review fields contain English language | No, 92% of negative, 94% of positive reviews are in English. |
| H01-06 | Negative and positive reviews are done in the same language per review | No, there are some reviews where different languages are use within one row of review |
| H01-07 | The reviews are distributed evenly over a given time frame | Yes, nearly uniformly distributed between 2015-08-04 and 2017-08-3 (2 years) |
| H01-08 | No duplicated reviews are contained within the data | No, 527 duplicates have been removed |
| H01-09 | There exists a well defined short lists of tags used | No, in total there are 2,428 different tags. However, the most frequently used 25 tags make up 87% of all tags. |
| H01-10 | The number of reviews per hotel strongly differs | Yes. There are hotels with nearly 5,000 reviews within the two years observed and hotels with less than 10 reveiews. |
| H01-11 | The review data contains only data of still existing and operating hotels | Yes, apart from one hotel, for which no recent (in 2017) reviews exist. All other hotels appear to have constantly received reviews during the observation period. |
| H01-12 | The reviews come from travellers all over the globe and the review data set is unbiased with respect to a reviewer nationality | No, 48% of reviews are given by British citizen, hence the United Kingdom is heavily over-represented |
| H01-13 | Most reviewers have in total only provided a few reviews | Yes, most reviewers have provided only a single review, 64% of reviewers have provided 5 or less reviews.


# Inits
Initialisations and loading of libraries required:
```{r, warning = FALSE, message = FALSE, results = FALSE}
rm(list=ls())   # clear all
graphics.off()  # close all
Sys.setenv(LANG = "en")
# Libs
# library(checkpoint) # Package management to avoid incompatibilities; not required when dockerized
# checkpoint('2017-10-01') # Use all packages as of date provided
library(data.table) # Efficient and fast data munging in R; just do it
options(datatable.auto.index=FALSE) # Bug fixing for v < 1.9.5
library(ggplot2) # Good quality plots
library(rprojroot) # Rroot folder management
ROOT <- find_root(has_file("PROJECT_ROOT_DIR"))
library(knitr) # Knitting R markdown reports
library(skimr) # Nice initial data overview
library(DataExplorer) # Nice summary statistics
library(leaflet) # Interacitve maps
library(sp) # Spatial data package
library(rworldmap) # Enables mapping of country level and gridded user datasets.
library(digest) # Hash computation
library(tidyr) # E.g. unnest data table for tags
library(stringr)
source(paste0(ROOT, "/02_code/lib.R"))
options(width = 100)
```

# Settings
All settings for this reports at a single spot here>
```{r}
s <- list() # List for storing settings
s$file_inf <- paste0(ROOT, '/01_data/02_unzipped/ID01/Hotel_Reviews.csv') # Data file name for challenge
s$file_kag <- paste0(ROOT, '/01_data/02_unzipped/ID02/Hotel_Reviews.csv') # Data file name form Kaggle
```

# Load data
```{r}
# Load inf data from csv
dt_inf <- fread(s$file_inf)
# Load Kaggle data from csv
dt_kag <- fread(s$file_kag)
```

# Compare data sets
```{r}
dim(dt_inf)
dim(dt_kag)
# Same dimensions

# Identical?
all.equal(dt_inf, dt_kag)
# Yes, H1 correct.
```

Cleanup as data is identical.
```{r}
rev <- dt_inf 
rm(list = c('dt_kag', 'dt_inf'))
```

# Data type conversions
```{r}
# What re the coltypes
col_types(rev)
```

Loading via `data.table::fread` coerced nearly perfect, just date is required to be coerced.

## Review_Date
```{r}
rev$Review_Date[1:10]
# Assume Review_Date is in format mm/dd/yyyy
rev[!grepl(pattern = '^\\d{1,2}\\/\\d{1,2}\\/\\d{4}', x = Review_Date)] # Yes
# Change data type to date
rev[, Review_Date := as.Date(Review_Date, format = '%m/%d/%Y')]
# Check
rev$Review_Date[1:5]
# Any NA created?
rev[is.na(Review_Date)] # No
```

## Review texts
Convert to lower case.
```{r}
rev[, Negative_Review := tolower(Negative_Review)]
rev[, Positive_Review := tolower(Positive_Review)]
```



# Exploratary data analysis
Exploratory data analysis is the process to get to know the data, so that hypothesis can be generated and tested.

## Initial inspection of data
```{r}
# Inspect first line
knitr::kable(rev[1])

# Quick overview
skim_with(numeric = list(hist = NULL),
          integer = list(hist = NULL)) # Do not show histogram as this is not displayed nicely within R markdown
skim(rev)

# Get introduced to your dataset using the DataExplorer package
introduce(rev) 
plot_intro(rev)
plot_missing(rev)
```

## Visualize distributions for all continuous features:

```{r}
plot_histogram(rev)
```


# Amend dataset

## Country
```{r}
rev[!is.na(lat), country := coords2country(cbind(lng, lat))]
# Inspect
rev[, .N, by = country]
```


# Hotel master data

## Generate master data
```{r}
# Look at column names
names(rev)
# Pick hotel related columns
cols_hot <- c('Hotel_Name', 'Hotel_Address', 'country', 'lat', 'lng', 'Total_Number_of_Reviews')
# Inspect
rev[1:3, cols_hot, with = F]
# Take only unique entries
hot <- unique(rev[, cols_hot, with = F])
hot[1:5]
dim(hot)
# Is combination of name and address uid?
sum(duplicated(hot[, .(Hotel_Name, Hotel_Address)])) # yes.
```

Note, that there where 1,493 unique hotel addresses, 1,492 unique hotel names and now 1,494 unique combinations of name, address and geolocation.  This implies, that there are still a few duplicates in the master data.

## Check hotel name
Note that there are names with only two letters. Check:

```{r}
hot[, n_name := nchar(Hotel_Name)]
hot[n_name <= 7] # Inspect names with less than 8 characters
```

Looks ok, aparat from name `41`

Inspect duplicates
```{r}
(dup_name <- unique(hot[duplicated(Hotel_Name), Hotel_Name]) )
hot[Hotel_Name %in% dup_name] # Looks ok, just same name
```

## Check hotel address
```{r}
(dup_addr <- unique(hot[duplicated(Hotel_Address), Hotel_Address]) )
hot[Hotel_Address %in% dup_addr] # Looks ok, maybe changed name or same building
```

## Check geolocations
```{r}
hot[is.na(lat) | is.na(lng)]
# 17 hotels with missing geolocations
```


## Is average hotel score unique for given hotel?
```{r}
dim(hot)
dim(unique(rev[, c(cols_hot, 'Average_Score'), with = F]))
```

Yes, so take average score as part of master data.
```{r}
hot <- unique(rev[, c(cols_hot, 'Average_Score'), with = F])
```

## Generate hotel id
```{r}
hot <- cbind(data.table(id_hot = 1:nrow(hot)), hot)
# Also add to review data
rev <- merge(rev, 
             hot[, .(id_hot, Hotel_Name, Hotel_Address)],
             by = c("Hotel_Name", "Hotel_Address"),
             all.x = T)
```



## Interactive visualization of hotel locations
Kindly borrowed from [Jason Liu](https://www.kaggle.com/jiashenliu/quick-visualization-of-data) using the awesome `leaflet` package.
```{r results='asis' ,message=FALSE, warning=FALSE}
leaflet(data = hot) %>% 
  addProviderTiles(providers$Stamen.TonerLite) %>% 
  addMarkers(popup = ~Hotel_Address, clusterOptions = markerClusterOptions())
```

There appear to be only dedicated locations:

* London
* Amsterdam
* Paris
* Milan
* Barcelona
* Vienna


## Add city column and city center geolocations

```{r}
# Inspect
hot[, .N, by = country]
# Map cities and geolocations (source https://www.latlong.net)
hot[country == "United Kingdom", c('city', 'city_lat', 'city_lng') := list('London', 51.509865, -0.118092)]
hot[country == "Netherlands", c('city', 'city_lat', 'city_lng') := list('Amsterdam', 52.37403, 4.88969)]
hot[country == "France", c('city', 'city_lat', 'city_lng') := list('Paris', 48.8566, 2.3522)]
hot[country == "Spain", c('city', 'city_lat', 'city_lng') := list('Barcelona', 41.3896, 2.1624)]
hot[country == "Italy", c('city', 'city_lat', 'city_lng') := list('Milan', 45.46427, 9.18951)]
hot[country == "Austria", c('city', 'city_lat', 'city_lng') := list('Vienna', 48.210033, 16.363449)]
```


## Add number of reviews per hotel within dataset
```{r}
# Compute and add number of reviews column
hot <- merge(hot,
             rev[, .(n_reviews = .N), by = id_hot],
             by = 'id_hot')
```




# Inspect review data

## Review language 
### Detect language of negative review texts
```{r}
# Negative reviews
rev[, lang_neg_rev := cld2::detect_language(Negative_Review)] 
rev[, .N, by = lang_neg_rev][order(N, decreasing = T)]
# Inspect
rev[lang_neg_rev == "de", Negative_Review] # Looks alright!
```

How much in English?
```{r}
rev[lang_neg_rev == "en", .N] / nrow(rev) # 92%
```



### Detect language of positive review texts
```{r}
rev[, lang_pos_rev := cld2::detect_language(Positive_Review)]
```

How much in English?
```{r}
rev[lang_pos_rev == "en", .N] / nrow(rev) # 94%
```



### Reviews in different languages
```{r}
# How many?
rev[lang_neg_rev == "en" & lang_pos_rev != "en", .N]
# Inspect
rev[lang_neg_rev == "en" & lang_pos_rev != "en", .(Negative_Review, Positive_Review)][1]
```


## Time of review
```{r}
ggplot(rev) + geom_density(aes(x = Review_Date))
rev[, min(Review_Date)]
rev[, max(Review_Date)]
```

More or less uniquely distributed over time

## Duplicate reviews
Are there any duplicates?
```{r}
sum(duplicated(rev)) # Yes! 527
# Create hash per row (could simply use duplicated function but so easy to inspect and UID generated)
cols <- names(rev)
rev$input_hash <- apply(rev[ , cols , with = F] , 1 , paste , collapse = "-" )
rev$hash <- apply(rev[ , 'input_hash' , with = F] , 1 , digest , algo = "md5" )
rev$input_hash <- NULL
# Find duplicate hashes
dup_hash <- rev[duplicated(rev$hash), unique(hash)]
length(dup_hash) # 527
# Inspect example
rev[hash == dup_hash[1]]
# Remove dups
rev <- rev[!duplicated(rev)]
# Use hash as id column: rename
setnames(rev, 'hash', 'id_rev')
```

## Plausicheck word counts
```{r}
# Inspect
rev[, .N, by = Review_Total_Negative_Word_Counts][order(N)]
rev[Review_Total_Negative_Word_Counts == 0, .(Review_Total_Negative_Word_Counts, Negative_Review)][1:3]
rev[Review_Total_Negative_Word_Counts == 2, .(Review_Total_Negative_Word_Counts, Negative_Review)][1:3]
rev[Review_Total_Negative_Word_Counts == 3, .(Review_Total_Negative_Word_Counts, Negative_Review)][1:3]
rev[Review_Total_Negative_Word_Counts == 4, .(Review_Total_Negative_Word_Counts, Negative_Review)][1:3]
# Roughly right, but not exact. is 'no negative' default value?
rev[, .N, by = Negative_Review][order(N, decreasing = T)][1:10]
rev[is.na(Negative_Review), .N]
# What about positive?
rev[, .N, by = Positive_Review][order(N, decreasing = T)][1:10]
```


## Inspect hotels with most and fewest reviews respectively
```{r}
# How many reviews per hotel?
rev[, .N, by = id_hot][order(N)]
# Inspect
# Inspect reviews for hotes with many and few reviews
ggplot(data = rev[id_hot %in% c(1019, 1383, 187, 487)],
       aes(x = Review_Date, y = Reviewer_Score)) +
  geom_point() + 
  facet_grid("id_hot") + 
  ggtitle("Top two hotels with most and fewest reviews respectively")
```

## Are there hotels which are no longer open?

Generate stat data for reviews over time
```{r}
# Aggregate and join together
stat_rev <- merge(
  hot[, .(id_hot)],
  rev[Review_Date < '2016-01-01', .(n_b4_2016 = .N), by = id_hot],
  by = 'id_hot',
  all.x = T)
stat_rev <- merge(
  stat_rev,
  rev[Review_Date > '2017-01-01', .(n_after_2017 = .N), by = id_hot],
  by = 'id_hot',
  all.x = T)
# Any NAs?
inspect_na(stat_rev) # Yes, set to zero.
stat_rev[is.na(n_b4_2016), n_b4_2016 := 0]
stat_rev[is.na(n_after_2017), n_after_2017 := 0]
inspect_na(stat_rev)
```

Inspect hotels, where there are no recent reviews
```{r}
stat_rev[n_after_2017 == 0] # Only two hotels, inspect
ggplot(data = rev[id_hot %in% c(7, 80)],
       aes(x = Review_Date, y = Reviewer_Score)) +
  geom_point() + 
  facet_grid("id_hot")
```

The hotel with id 7 seems indeed to have no recent reviews, whilst the hotel with id 80 simply appears to have very few reviews and is not necessarily closed at the end of the observed time range.



# Tag data

## Extract tag data from review data
Tags are given as nested elements with a 1:n relationship for review and tag. Generate data table in long format.

```{r}
# Split tag with respect to ','
rev[, tag_split := strsplit(rev[, Tags], ',')]
# Add number of tags to review data table
rev[, n_tags := str_count(Tags, ',') + 1]
# How many tags per review?
rev[, .N, by = n_tags][order(N)]
# Extract nested list to data table in long format
tags <- unnest(rev[, .(id_rev, tag_split)])
# Inspect 
print(tags)
# Clean up 
setnames(tags, 'tag_split', 'tag') # Rename column
tags[, tag := gsub(pattern = ']', replacement = '', x = tag, fixed = T)] # Remove characters
tags[, tag := gsub(pattern = '[', replacement = '', x = tag, fixed = T)] # Remove characters
tags[, tag := gsub(pattern = "'", replacement = '', x = tag, fixed = T)] # Remove characters
tags[, tag := trim(tag)] # Remove leading and trailing blanks
tags[, tag := tolower(tag)]
# Inspect
print(tags)
```

## Inspect most frequent tag data

Create stats table for tag data.
```{r}
# Inspect count
(tag_stats <- tags[, .N, by = tag][order(N, decreasing = T)]) # 2428 different tags
# Look at most frequent ones
tag_stats[1:20]
tag_stats[N != 1]
# Look at frequency distribution
ggplot(tag_stats) + geom_point(aes(x = 1:nrow(tag_stats), y = N))
# Find ellbow
ggplot(tag_stats[N >= 300]) + geom_point(aes(x = 1:nrow(tag_stats[N >= 300]), y = N))
# Ellbow around 25
tag_stats[1:25] # Look at top tags
tags_most_freq <- tag_stats[1:25, tag]
# How much percentage covered with these tags
tags[tag %in% tags_most_freq, .N] / nrow(tags) # The top 25 tags cover 87% of tags, pareto
```

## Inspect nights stayed tags
```{r}
tag_stats[grepl(pattern = "stay", x = tag)]
# Inspect top 25 without stay tags
tag_stats[!grepl(pattern = "stay", x = tag), tag][1:25]
```

## Inspect room tags
```{r}
tag_stats[grepl(pattern = "room", x = tag)]
# Most tags are concerned with the room
tag_stats[grepl(pattern = "room", x = tag)][1:50]
```

## Inspect tags other than room or stays
```{r}
# Inspect remaining tags
tag_stats[!grepl(pattern = "room", x = tag) &
           !grepl(pattern = "stay", x = tag), 
          ][1:25]
```

# Reviewer data
The only attributes of the guest within data are:

* Total_Number_of_Reviews_Reviewer_Has_Given: Number of Reviews the reviewers has given in the past.
* Reviewer_Nationality

Inspect Total_Number_of_Reviews_Reviewer_Has_Given
```{r}
rev[, .N, by = Total_Number_of_Reviews_Reviewer_Has_Given][order(N)]
ggplot(rev) + 
  geom_histogram(aes(x = Total_Number_of_Reviews_Reviewer_Has_Given))
# How many percentage of reveiwers have given less or equal to 5 reviews?
rev[Total_Number_of_Reviews_Reviewer_Has_Given %in% 1:5, .N] / nrow(rev)
```

Inspect Reviewer_Nationality
```{r}
rev[, .N, by = Reviewer_Nationality][order(N, decreasing = T)]
rev[, .N, by = Reviewer_Nationality][order(N, decreasing = T)][1:20]
# Perdentage of British
rev[Reviewer_Nationality == 'United Kingdom', .N] / nrow(rev)
```




# Save preprocessed data
```{r}
saveRDS(rev,  paste0(ROOT, '/01_data/03_preprocessed/reviews.rds'))
saveRDS(hot,  paste0(ROOT, '/01_data/03_preprocessed/hotels.rds'))
saveRDS(tags,  paste0(ROOT, '/01_data/03_preprocessed/tags.rds'))
saveRDS(tag_stats,  paste0(ROOT, '/01_data/03_preprocessed/tag_stats.rds'))
```

