---
title: "05 - Feature engineering"
author: "Jens Linden"
output:
  html_document:
    number_sections: yes
    toc: yes
---

STATUS: Development

# Management summary

Create analytics base table for hotels and reviews. 

The following features have been selected or created:

* Hotel
    - `dist_city_centre`: Distance of hotel to city centre in meters
    - `Total_Number_of_Reviews`: Number of reviews given to hotel as a proxy for the size of the hotel
    - `city`: City of hotel

* Review
    - `month`: Month when review was given
    - `n_nights`: Number of nights the reviewer stayed in the hotel
    - `trip_type`: Whether `business trip` or `leisure trip`. `NA`s set to `unknown` when info missing.
    - `is_mobile_device`: Whether the trip was booked via a mobile device or not (`yes` or `no`)
    - `group_type`: Whether the reviewer travelled alone or in a group. Possible values are `couple`, `solo traveler`,`group`, `family with young children`, `family with older children`, `travelers with friends`, `with a pet`
    - `is_superior_room`: Whether a 'superior' room was booked (`yes` or `no`)
    - `room_size`: Size of room booked (`single`, `double` or `family`. `NA`s set to `unknown` when info missing.)

# Inits
Initialisations and loading of libraries required:
```{r, warning = FALSE, message = FALSE, results = FALSE}
rm(list=ls())   # clear all
graphics.off()  # close all
Sys.setenv(LANG = "en")
# Libs
# library(checkpoint) # Package management to avoid incompatibilities; not required when dockerized
# checkpoint('2017-10-01') # Use all packages as of date provided
library(data.table) # Efficient and fast data munging in R; just do it
options(datatable.auto.index=FALSE) # Bug fixing for v < 1.9.5
library(ggplot2) # Good quality plots
library(rprojroot) # Rroot folder management
ROOT <- find_root(has_file("PROJECT_ROOT_DIR"))
library(geosphere) # Compute distance of geospatial data
library(knitr) # Knitting R markdown reports
library(tidyr) # extract numeric from string
library(tm)
source(paste0(ROOT, "/02_code/lib.R"))
options(width = 100)
```

# Load data
```{r}
(hot <- readRDS(paste0(ROOT, '/01_data/03_preprocessed/hotels.rds')))
(rev <- readRDS(paste0(ROOT, '/01_data/03_preprocessed/reviews.rds')))
(tag <- readRDS(paste0(ROOT, '/01_data/03_preprocessed/tags.rds')))
```

# Hotel

## Distance of hotel from city centre
```{r}
dist_ <-  distm(hot[, .(lng, lat)], hot[, .(city_lng, city_lat)], fun = distHaversine)
hot[, dist_city_centre := diag(dist_)]
# Inspect
hot[, .(id_hot, Hotel_Address, dist_city_centre)]
ggplot(hot[!is.na(dist_city_centre)]) + 
  geom_histogram(aes(x = dist_city_centre), bins = 50) +
  facet_grid("city") # Looks plausible
```

## Size of hotel
Use proxy as number of total reviews given `Total_Number_of_Reviews`.
```{r}
# Inspect
ggplot(hot) + geom_histogram(aes(x = Total_Number_of_Reviews))
```

## City
As country and city are identical, choose city
```{r}
hot[, .N, by = city]
```

## Build analytics base table (ABT) for hotels
```{r}
abt_hot <- hot[, .(id_hot, dist_city_centre, Total_Number_of_Reviews, city)]
abt_hot <- abt_hot[complete.cases(abt_hot)]
print(abt_hot)
```

## Open points
The following features should be considered in the future to obtain comparable cluster of hotels:

* Stars: Category of hotel (e.g. 5 stars) is already a comparison indicator and should be utilized to determine peer hotels
* Price: Average room price of hotel


# Reviews

## Initialise ABT with month
```{r}
abt_rev <- rev[, .(id_rev, id_hot, month = format(Review_Date,"%B"))]
```

## From tags

Get stats
```{r}
(tag_stats <- tag[, .N, by = tag][order(N, decreasing = T)])
```


### Number of stayed nights
```{r}
# Inspect
tag_stats[grepl(pattern = "stayed", x = tag)]
# Extract number of nights
tag[grepl(pattern = "stayed", x = tag), 
    n_nights := extract_numeric(tag)]
# Plausicheck
print(tag) # Visual
tag[!is.na(n_nights)][, sum(duplicated(id_rev))] # ok, only one night tag for each review
# Add to abt
(abt_rev <- merge(abt_rev,
                 tag[!is.na(n_nights), .(id_rev, n_nights)],
                 by = 'id_rev',
                 all.x = T))
```

Cleanup
```{r}
# Remove used tags
tag <- tag[!grepl(pattern = "stayed", x = tag)]
tag$n_nights <- NULL
# Recompute tag statistics
(tag_stats <- tag[, .N, by = tag][order(N, decreasing = T)])
```


### Trip type
Whether leisure or business
```{r}
# Check tag labels
tag_stats[!grepl(pattern = "triple", x = tag) & grepl(pattern = "trip", x = tag)]
# Plausicheck
tag[tag %in% c('leisure trip', 'business trip')][, sum(duplicated(id_rev))] # ok, only one type for each review
# Add to abt
(abt_rev <- merge(abt_rev,
                 tag[tag %in% c('leisure trip', 'business trip'), .(id_rev, trip_type = tag)],
                 by = 'id_rev',
                 all.x = T))
# Set NA to 'unknown'
abt_rev[is.na(trip_type)]
abt_rev[is.na(trip_type), trip_type := 'unknown']
# Inspect
print(abt_rev[1:3])
```

Cleanup
```{r}
# Remove tags used
tag <- tag[tag %ni% c('leisure trip', 'business trip')]
# Recompute tag statistics
(tag_stats <- tag[, .N, by = tag][order(N, decreasing = T)])
```

### Mobile device booker
```{r}
# Check tag labels
tag_stats[grepl(pattern = "device", x = tag) ]
tag[grepl(pattern = "device", x = tag)][, sum(duplicated(id_rev))]
# Add to abt
ids_mobile <- tag[tag %in% c('submitted from a mobile device'), id_rev]
abt_rev[id_rev %in% ids_mobile, is_mobile_device := 'yes']
abt_rev[id_rev %ni% ids_mobile, is_mobile_device := 'no']
# Inspect
abt_rev[1:3]
abt_rev[, .N, by = is_mobile_device]
```

Cleanup
```{r}
# Remove tags used
tag <- tag[tag %ni% c('submitted from a mobile device')]
# Recompute tag statistics
(tag_stats <- tag[, .N, by = tag][order(N, decreasing = T)])
```

### Group type
```{r}
# Show remaining tags, which do not contain room
tag_stats[!grepl(pattern = "room", x = tag), ][1:20]
# Define relevant tags for group type
tags_group <- c('couple',
               'solo traveler',
               'group',
               'family with young children',
               'family with older children',
               'travelers with friends',
               'with a pet')
# Check duplicates
tag[tag %in% tags_group][, sum(duplicated(id_rev))]
# Yes, multiple choices are possible. However, only a few, so ignore
(abt_rev <- merge(abt_rev,
                 tag[tag %in% tags_group][!duplicated(id_rev)][, .(id_rev, group_type = tag)],
                 by = 'id_rev',
                 all.x = T))
# Any NAs?
abt_rev[is.na(group_type)] # No, good! Group type was specified for each review.
```

Cleanup
```{r}
# Remove tags used
tag <- tag[tag %ni% tags_group]
rm(tags_group)
# Recompute tag statistics
(tag_stats <- tag[, .N, by = tag][order(N, decreasing = T)])
```

### Superior room
```{r}
# Inspect tags indicating superior
tag_stats[grepl(pattern = "superior", x = tag) |
           grepl(pattern = "deluxe", x = tag) , ]
tag_stats[grepl(pattern = "executive", x = tag) , ]
tag_stats[grepl(pattern = "king room", x = tag) |
           grepl(pattern = "queen room", x = tag) , ]
tag_stats[grepl(pattern = "king suite", x = tag) |
           grepl(pattern = "queen suite", x = tag) , ]
# Define flag for superior
tag[grepl(pattern = "superior", x = tag) |
      grepl(pattern = "deluxe", x = tag) |
      grepl(pattern = "executive", x = tag) |
      grepl(pattern = "king room", x = tag) |
      grepl(pattern = "queen room", x = tag) |
      grepl(pattern = "king suite", x = tag) |
      grepl(pattern = "queen suite", x = tag),
    is_superior_room := 'yes']
print(tag)
# Check
tag[is_superior_room == 'yes', sum(duplicated(id_rev))] # good, no dups
# Add to abt
(abt_rev <- merge(abt_rev,
                 tag[is_superior_room == 'yes', .(id_rev, is_superior_room)],
                 by = 'id_rev',
                 all.x = T))
abt_rev[is.na(is_superior_room), is_superior_room := 'no']
abt_rev[, .N, by = is_superior_room]
```

Cleanup
```{r}
tag$is_superior_room <- NULL
```

### Room size
```{r}
# Inspect
tag_stats[1:50]
# Double rooms
(tags_double <- tag_stats[grepl(pattern = "double", x = tag) |
                           grepl(pattern = "twin", x = tag), tag])
tag[tag %in% tags_double, room_size := 'double']
# Single rooms 
tag[grepl(pattern = "single room", x = tag) |
      grepl(pattern = "single suite", x = tag), room_size := 'single']
# Family rooms
tag_stats[grepl(pattern = "family", x = tag)]
tag[grepl(pattern = "family", x = tag), room_size := 'family']
# Check dups
tag[!is.na(room_size), sum(duplicated(id_rev))] # No duplicates, good.
# Add to abt
(abt_rev <- merge(abt_rev,
                 tag[!is.na(room_size), .(id_rev, room_size)],
                 by = 'id_rev',
                 all.x = T))
abt_rev[is.na(room_size), room_size := 'unknown']
print(abt_rev)
# Plausichecks
abt_rev[, .N, by = room_size]
ids_unkn <- abt_rev[room_size == 'unknown', id_rev]
# Cross check in review data
rev[id_rev %in% ids_unkn, .(tag_split)] # Looks indeed, that from some review tags, the room size cannot be deduced

```

Cleanup
```{r}
rm(ids_unkn)
tag$room_size <- NULL
tag <- tag[!(grepl(pattern = "room", x = tag) |
      grepl(pattern = "suite", x = tag))]
# Recompute tag statistics
(tag_stats <- tag[, .N, by = tag][order(N, decreasing = T)])
tag_stats[1:50]
# Most tags processed, leave it here.
```

## From text reviews


### tst

```{r}
# # Create corpus
# corp <- Corpus(VectorSource(rev$Negative_Review))
# # Pre-Processing
# getTransformations()
# corp <- tm_map(corp, stripWhitespace) # removing extra spaces (keeping only single space)
# corp <- tm_map(corp, tolower) # transforming to lower case
# corp <- tm_map(corp, removeNumbers)
# corp <- tm_map(corp, stemDocument)
```



# Save data
```{r}
saveRDS(abt_hot,  paste0(ROOT, '/01_data/03_preprocessed/abt_hot.rds'))
saveRDS(abt_rev,  paste0(ROOT, '/01_data/03_preprocessed/abt_rev.rds'))
```

